---
title: "SIR Final Size"
author: "Loc Huynh"
date: '2023'
output:
  html_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
SIR model 
dS/dt = -beta SI
dI/dt = beta SI - nu I
dR/dt = nu I

We're going to use "dynamic allocation" of data structures

ODE question: for a given, N, beta, nu, I(0), what is the final epidemic size?

MC question: for the same parameters, what is the expected final epidemic size? E(R)

Numerical experiment question: What does the distribution of the final epidemic size look like? (Run lots of epidemics with identical parameters and report a histogram)

# The final size of the Stochastic SIR Model

```{r}

N = 1000 #S+I+R = 1000
num_steps = 2*N
betaN = 0.3 #beta*N*I = 50 means 50 infections per day 
beta = betaN/N 
nu = 0.2 #recovery time. The average is 1/nu. The average time is 50 days 
R0 = betaN/nu

num_epidemics = 1000 #how many time do you want it to run

final_epidemic_size = c()
for (i in 1:num_epidemics) {
  I = c()
  R = c()
  t = c()
  I = c(I,1) # concatenation
  R = c(R,0)
  t = c(t,0)
  n = 2
  while(n <= num_steps & I[n-1] > 0 & R[n-1]>= 0){
    infection_rate = beta*(N-I[n-1]-R[n-1])*I[n-1]
    recovery_rate = nu*I[n-1]
    tau = rexp(1, rate = infection_rate + recovery_rate)
    if (runif(1) < infection_rate/(infection_rate + recovery_rate)){
      I = c(I,I[n-1] + 1)
      R = c(R,R[n-1])
    } else {
      I = c(I,I[n-1] - 1)
      R = c(R,R[n-1] + 1)
    }
    t = c(t,t[n-1]+tau)
    n = n+1
  }
  final_epidemic_size = c(final_epidemic_size,max(R))
}

pdf("3stosircase4.pdf")

h = hist(final_epidemic_size,
     main = "Final epidemic size distribution",
     xlab = "Final epidemic size",
     freq = TRUE)

dev.off() 

long_idx = which(final_epidemic_size > 300) #return the indices of vec that satisfies

print(paste0("The mean of the epidemic size is ",mean(final_epidemic_size)))

print(paste0("The mean of the epidemic size is ",mean(final_epidemic_size[long_idx])))
#vec[long_idx]: the value that associated with the index


pdf("3stosircase42.pdf")

h_approx = hist(final_epidemic_size[long_idx], 
                 main = "Major Outbreak epidemic size distribution",
                 xlab = "Final epidemic size")
dev.off()
```

# Relationship between Final size of the epidemic in the determinstic solution and $R_0$
```{r}
R_0_vec = seq(from = 0, to = 5, length.out = 10000)
final_size = function(R_0){
  eqn = function(r_star){
    1 - r_star - exp(-R_0 * r_star)
  }
  result1 = uniroot(eqn, lower = 0, upper = 1)$root
  result2 = uniroot(eqn, lower = 1e-12, upper = 1, extendInt = "yes")$root
  
  result = max(result1, result2)
  return(result)
}

final_vec = as.vector(10000)
for (i in 1:10000){
  R_0 = R_0_vec[i]
  final_vec[i] <- final_size(R_0 = R_0)
}
pdf("2finalR0.pdf")

plot(R_0_vec, final_vec, type = "s", xlab = "R_0", ylab = "Final epidemic size")

dev.off()
```


