---
title: "SIR Final Size"
author: "Loc Huynh"
date: '2023'
output:
  html_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
SIR model 
dS/dt = -beta SI
dI/dt = beta SI - nu I
dR/dt = nu I

We're going to use "dynamic allocation" of data structures

ODE question: for a given, N, beta, nu, I(0), what is the final epidemic size?

MC question: for the same parameters, what is the expected final epidemic size? E(R)

Numerical experiment question: What does the distribution of the final epidemic size look like? (Run lots of epidemics with identical parameters and report a histogram)

#Solution of stochastic SIR
```{r}
#comment multiple lines: Crtl + Shift + C
N = 1000 #S+I+R = 1000
num_steps = 2*N
betaN = 0.3 #beta*N*I = 50 means 50 infections per day 
beta = betaN/N 
nu = 0.2 #recovery time. The average is 1/nu. The average time is 50 days 
R0 = betaN/nu

#final_epidemic_size = c()
#for (i in 1:num_epidemics) {
# PLACE epidemic simulation here  
#  final_epidemic_size = c(final_epidemic_size,max(R))
#}
#hist(final_epidemic_size)

I = c()
R = c()
t = c()
I = c(I,1) # concatenation
R = c(R,0)
t = c(t,0)
n = 2
current_R = c()
current_R = c(current_R,beta*(N-I[n-1]-R[n-1])/nu)
while(n <= num_steps & I[n-1] > 0 & R[n-1]>= 0){
  infection_rate = beta*(N-I[n-1]-R[n-1])*I[n-1]
  recovery_rate = nu*I[n-1]
  tau = rexp(1, rate = infection_rate + recovery_rate)
  current_R = c(current_R,beta*(N-I[n-1]-R[n-1])/nu)
  if (runif(1) < infection_rate/(infection_rate + recovery_rate)){
    I = c(I,I[n-1] + 1)
    R = c(R,R[n-1])
  } else {
    I = c(I,I[n-1] - 1)
    R = c(R,R[n-1] + 1)
  }
  t = c(t,t[n-1]+tau)
  n = n + 1
}
S = N - I - R

print(paste0("The basic reproduction number is ",R0))
print(paste0("The final epidemic size is ",max(R)))



# Opening the graphical device
pdf("3stosircase22.pdf")
#Plot them seperately
#plot(t,S,type="s")
#plot(t,I,type="s")
#plot(t,R,type="s")

#Plot them together
plot(t,S,type="s", col = 1, xlab = "Time", ylab = "", 
     xlim = c(0,max(t)), ylim= c(0,N))
lines(t,I,col = 2)
lines(t,R,col = 3)
legend(x = "topright", c("I(t)", "R(t)","S(t)"), col = c(2, 3, 1), lty = 1)
title("Plot of S(t), I(t) and R(t) ")

# Closing the graphical device
dev.off() 
```
#The deterministic solution of SIR model:
```{r, message=FALSE, warning=FALSE}
library(deSolve)

deter_sir = function(time, y, parms){
  I = y[1]
  R = y[2]
  my_beta = parms[1]
  my_nu = parms[2]
  pop = parms[3]
  
  return(list(c(
    I = my_beta*I*(pop - I - R) - my_nu*I,
    R = my_nu*I
  )))
}

time = seq(from = 0, to = 200, length.out = 10000)

sol = lsoda(
  y = c(1, 0), #Initial I and R
  parms = c(beta, nu, N),
  times = time,
  func = deter_sir
)


# Opening the graphical device
pdf("3detsircase2.pdf")

# Creating a plot
plot(time, sol[, 2], type = "l", lwd = 2, 
     col = 2, xlab = "time", ylab = "", xlim = c(0,200), ylim = c(0,1000))
lines(time, sol[, 3], col = 3)
lines(time, N-sol[,2]-sol[,3], col = 1)
legend(x = "topright", c("I(t)", "R(t)","S(t)"), col = c(2, 3, 1), lty = 1)
title("Plot of S(t), I(t) and R(t) ")

# Closing the graphical device
dev.off() 

print(paste0("The final epidemic size is ",max(sol[,3])))

```
